---
title: "AE 03 Flights and pipes"
subtitle: "Suggested answers"
---

::: callout-important
These are suggested answers.
This document should be used as a reference only; it's not designed to be an exhaustive key.
:::

# Introduction

Have you ever had a flight delayed?
How common is that?
Which airlines are delayed more than others?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **nycflights13** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(nycflights13)
```

## Data

The dataset we will explore is called `flights`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse-flights
glimpse(flights)
```

# Exploring flights

## Exercise 1

Fill in the blanks:

The `flights` data frame has `r nrow(flights)` rows.
Each row represents a **single flight departing NYC in 2013**.

## Exercise 2

What are the names of the variables in `flights`.

```{r}
#| label: column-names
# add code here
colnames(flights)
```

**Answer:**
`year, month, day, dep_time, sched_dep_time, dep_delay, arr_time, sched_arr_time, arr_delay, carrier, flight, tailnum, origin, dest, air_time, distance, hour, minute, time_hour`

## Exercise 3 - `select()`

-   Make a data frame that only contains the variables `dep_delay` and `arr_delay`.

```{r}
#| label: select-vars
# add code here
flights |> 
  select(dep_delay, arr_delay)
```


-   Make a data frame that keeps every variable except `dep_delay`.

```{r}
#| label: exclude-vars
# add code here
flights |> 
  select(!dep_delay)
```

-   Use the `select` helper `contains()` to make a data frame that includes the variables associated with the arrival, i.e., contains the string `"arr_"` in the name.

```{r}
#| label: arr-vars
# add code here
flights |> 
  select(contains("arr_"))
```

## Exercise 4 - `slice()`

-   Display the first five rows of the `flights` data frame.

```{r}
#| label: slice
# add code here
flights |> 
  slice_head(n = 5)
```

## Exercise 5 - `arrange()`

-   Let's arrange the data by departure delay, so the flights with the shortest departure delays will be at the top of the data frame.

```{r}
#| label: arrange-delays
# add code here
flights |> 
  arrange(dep_delay)
```

-   What does it mean for the `dep_delay` to have a negative value?

**Answer:** It means the flight departed earlier than scheduled.

-   Arrange the data by descending departure delay, so the flights with the longest departure delays will be at the top.

```{r}
#| label: arrange-delays-desc
# add code here
flights |> 
  arrange(desc(dep_delay))
```

-   Create a data frame that only includes the plane tail number (`tailnum`), carrier (`carrier`), and departure delay for the flight with the longest departure delay. What is the plane tail number (`tailnum`) for this flight?

```{r}
#| label: longest-delay
# add code here
# Store the new dataframe so we can reference it later in the narrative
# using inline code
most_delayed <- flights |>
  select(tailnum, carrier, dep_delay) |>
  slice_max(dep_delay)
# Display the resulting dataframe
most_delayed
```

**Answer:** The most delayed flight had tail number `r most_delayed$tailnum`.

## Exercise 6 - `filter()`

-   Filter for all rows where the destination airport is RDU.

```{r}
#| label: rdu
# add code here
flights |> 
  filter(dest == "RDU")
```

-   Filter for all rows where the destination airport is RDU and the arrival delay is less than 0.

```{r}
#| label: rdu-ontime
# add code here
flights |> 
  filter(dest == "RDU", arr_delay < 0)
```

-   Describe what the code is doing in words.

**Answer:** The following code keeps flights going to airports RDU or GSO, where the arrival delay or the departure delay are less than 0. In other words, flights to RDU or GSO that departed early or arrived early.

```{r}
#| label: nc-early
flights |>
  filter(
    dest %in% c("RDU", "GSO"),
    arr_delay < 0 | dep_delay < 0
  )
```

**Hint:** Logical operators in R:

| operator      | definition                                               |
|:--------------|:---------------------------------------------------------|
| `<`           | is less than?                                            |
| `<=`          | is less than or equal to?                                |
| `>`           | is greater than?                                         |
| `>=`          | is greater than or equal to?                             |
| `==`          | is exactly equal to?                                     |
| `!=`          | is not equal to?                                         |
| `x & y`       | is x AND y?                                              |
| `x | y`       | is x OR y?                                               |
| `is.na(x)`    | is x NA?                                                 |
| `!is.na(x)`   | is x not NA?                                             |
| `x %in% y`    | is x in y?                                               |
| `!(x %in% y)` | is x not in y?                                           |
| `!x`          | is not x? (only makes sense if `x` is `TRUE` or `FALSE`) |

## Exercise 7 - `count()`

-   Create a frequency table of the destination locations for flights from New York's 3 major airports (EWR, LGA, JFK).

```{r}
#| label: count-dest
# add code here
flights |>
  filter(origin %in% c("EWR", "LGA", "JFK")) |>
  count(dest, sort = TRUE)
```

-   In which month was there the fewest number of flights? How many flights were there in that month?

```{r}
#| label: count-month
# add code here
flights |> 
  count(month, sort = TRUE)
```

**Answer:** The month with the fewest number of flights was February, which had 24,951 flights.

-   On which date (month + day) was there the largest number of flights? How many flights were there on that day?

```{r}
#| label: count-date
# add code here
flights |> 
  count(month, day, sort = TRUE)
```

**Answer:** The date with the largest number of flights was 11/27, with 1,014 flights. The large number of flights was likely due to Thanksgiving holiday travel.

## Exercise 8 - `mutate()`

-   Convert `air_time` (minutes in the air) to hours and then create a new variable, `mph`, the miles per hour of the flight.

```{r}
#| label: calculate-mph
# add code here
flights |> 
  mutate(
    air_time = air_time / 60,
    mph = distance / air_time
  )
```
