---
title: "AE 07: Survey types"
subtitle: Suggested answers
categories: 
  - Application exercise
  - Answers
---

::: callout-important
These are suggested answers.
This document should be used as a reference only; it's not designed to be an exhaustive key.
:::

# Getting started

## Packages

We'll use the **tidyverse** package for this analysis.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(readxl)
```

## Data

The data are available in the `data` folder.

```{r}
#| label: load-data
#| message: false
#| error: true
survey_raw <- read_excel("data/survey-2026-10-05.xlsx")
survey_raw
```

# Prep

Clean the data such that it has two variables:

1. `year`: The class of the students with levels `First-year`, `Sophomore`, `Junior`, and `Senior`, in that order.
2. `n_mon_classes`: The number of classes each student has on Mondays.

Save the resulting data frame as `survey`.

```{r}
#| label: clean-data
# Add your code here
survey <- survey_raw |>
  mutate(
    year = fct_relevel(year, "First-year", "Sophomore", "Junior", "Senior")
  )
survey
```


# Explore

Explore the relationship between `year` and `n_mon_classes` using a table and a plot.
Describe the relationship or the lack thereof .

```{r}
#| label: explore-data
# Add your code here
survey |>
  group_by(year) |>
  summarize(
    mean_n_classes = mean(n_mon_classes),
    med_n_classes = median(n_mon_classes),
    sd_n_classes = sd(n_mon_classes)
    )

survey |>
  ggplot(aes(x = year, y = n_mon_classes)) +
  geom_boxplot() 

survey |>
  ggplot(aes(x = n_mon_classes, color = year)) +
  geom_density()

survey |>
  ggplot(aes(x = year, fill = factor(n_mon_classes))) +
  geom_bar(position = "fill")

survey |>
  ggplot(aes(x = n_mon_classes, fill = year)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~year)
```
