---
title: Lab 3
subtitle: JOINing the Olympic Gymnastics Committee
categories: Lab
draft: false
---

## Learning objectives

By the end of the lab, you will...

-   Combine two data frames with `left_join()`, `inner_join()`, and `anti_join()`
-   Specify join keys with `join_by()` when the key columns have different names
-   Reshape data with `pivot_longer()` and `pivot_wider()`
-   Convert a character variable to a factor and reorder its levels with `fct_relevel()`
-   Continue developing your data wrangling and visualization skills with the tidyverse
-   Get more experience with the data science workflow using R, RStudio, Git, and GitHub

## Getting started

Log in to RStudio, clone your `lab-3` repo from GitHub, open your `lab-3.qmd` document, and get started!

::: {.callout-tip collapse="true"}
## Click here if you prefer to see step-by-step instructions

### Step 1: Login to Posit Cloud

-   Go to <https://posit.cloud> and go to our course workspace.

### Step 2: Clone the repo & start new RStudio project

-   Go to the GitHub course organization and click on the repo with the prefix **lab-3**.
    It contains the starter documents you need to complete the lab.

-   Click on the green **CODE** button and select **HTTPS** (it should already be selected by default).
    Click on the clipboard icon to copy the repo URL.

-   In RStudio, go to New Project ➛ New Project from Git Repository ➛ Git
-   Paste the URL of your assignment repo into the dialog box URL of your Git Repository.
-   Click OK. The repo will be cloned and an RStudio project will be created.
-   Go to the Files pane and click **lab-3.qmd** to open the template Quarto file.
    This is where you will write up your code and narrative for the lab.

### Step 3: Update the YAML

In `lab-3.qmd`, update the `author` field to your name, render your document, and examine the changes.
Then, in the Git pane, stage your changes, add a commit message (e.g., "Added author name"), and click **Commit**.
Then, push the changes to your GitHub repository and, in your browser, confirm that these changes have indeed propagated to your repository.

::: callout-important
If you run into any issues with the first steps outlined above, seek help before proceeding.
:::
:::

::: callout-important
Your repo also contains a `data/` folder and a `tests/` folder.
**Do not move, rename, edit, or delete them.**
The template reads data from `data/` and uses `tests/` to give you instant feedback on your answers (see "Checking your work" below).
:::

## Packages

In this lab we will work with the **tidyverse** package, which is a collection of packages for doing data analysis in a "tidy" way.

```{r}
#| eval: true
#| message: false
library(tidyverse)
```

-   **Run** the code cell by clicking on the green triangle (play) button for the code cell labeled `load-packages`. This loads the package to make its features (the functions and datasets in it) be accessible from your *Console*.
-   Then, **render** the document which loads this package to make its features (the functions and datasets in it) be available for other code cells in your Quarto document.

## Checking your work

This lab is **autograded**.
That changes two things about how you work:

1.  **Answers go into named objects.**
    Each question below tells you exactly what to name your answer (e.g., "save the result as `scores_codes`"), and the template's code cells already contain placeholders such as `scores_codes <- NULL` with a short comment describing what goes there.
    Replace the `NULL` (or `NA`) with your code, but **do not change the object names**.
    The autograder looks for these exact names; an answer stored under a different name receives no credit, even if the code is correct.

2.  **You can check your answers as you go.**
    After each question, the template contains a check chunk that runs `ottr::check()` on that question.
    Run it to see the results of the **public checks**: quick tests that confirm your answer has the right structure (correct columns, plausible number of rows, and so on).
    These checks are designed to tell you whether you're on the right track *without* revealing the answer.

::: callout-note
## Multiple choice and short answers

Some questions ask you to record an answer by assigning a value to a variable.
Letters and names must be typed as strings (in quotes), for example `q1_cause <- "B"` or `q9_highest <- "Antarctica"`.
Names should be spelled exactly as they appear in the data.
Numeric answers are typed as numbers without quotes, for example `n_routines <- 100`.
:::

::: callout-warning
Passing all public checks does **not** guarantee full credit.
When your submission is graded, additional **hidden tests** check your answers for full correctness.
An answer with the right shape but wrong values can pass every public check and still lose points, so always sanity-check your results yourself.
A check chunk that errors with something like "object not found" usually means you haven't run the cells above it -- run all chunks in order (or render the document) and try again.
:::

## Guidelines

As we've discussed in lecture, your plots should include an informative title, axes and legends should have human-readable labels, and careful consideration should be given to aesthetic choices.

Additionally, code should follow the [tidyverse style](https://style.tidyverse.org/).
Particularly,

-   there should be spaces before and line breaks after each `+` when building a `ggplot`,

-   there should also be spaces before and line breaks after each `|>` in a data transformation pipeline,

-   code should be properly indented,

-   there should be spaces around `=` signs and spaces after commas.

Furthermore, all code should be visible in the PDF output, i.e., should not run off the page on the PDF.
Long lines that run off the page should be split across multiple lines with line breaks.[^1]

[^1]: Remember, haikus not novellas when writing code!

::: callout-important
You should have at least 3 commits with meaningful commit messages by the end of the assignment.
:::

# Questions

## Part 1

**Bringing two data frames together**

In honor of National Gymnastics Day, you will work with real gymnastics data from the 2024 Paris Olympics!
For Part 1, you will work with two small data frames and join them with various join functions.
Both are created for you in the template with the code below:

-   `paris_scores`: Team final balance beam and floor exercise scores for four gymnasts.

```{r}
#| eval: true
paris_scores <- tribble(
  ~gymnast         , ~balance_beam , ~floor_exercise ,
  "Simone Biles"   , 14.40         , 14.70           ,
  "Sunisa Lee"     , 14.60         , 13.50           ,
  "Rebeca Andrade" , 14.10         , 14.20           ,
  "Jordan Chiles"  , 12.70         , 14.00
)

paris_scores
```

-   `noc_codes`: 3-letter National Olympic Committee (NOC) code of the country each of four gymnasts competed for.

```{r}
#| eval: true
noc_codes <- tribble(
  ~gymnast_name    , ~country_code ,
  "Simone Biles"   , "USA"         ,
  "Rebeca Andrade" , "BRA"         ,
  "Kaylia Nemour"  , "ALG"         ,
  "Qiyuan Qiu"     , "CHN"
)

noc_codes
```

::: callout-note
The `tribble()` function is helpful for creating small data frames (`tibble`s) with an easier to read row-by-row layout.
:::

### Question 1

Your friend writes the following code to join the `paris_scores` and `noc_codes` data frames and gets an error message:

```{r}
#| eval: false
paris_scores |>
  left_join(noc_codes)
```

```
Error in `left_join()`:
! `by` must be supplied when `x` and `y` have no common variables.
ℹ Use `cross_join()` to perform a cross-join.
```

**1a.**
Record in `q1_cause` the option that best explains why this error occurs:

   A. The two data frames have no column with the same name, so R cannot guess which columns to match rows on.

   B. The two data frames have different numbers of rows, so they cannot be joined.

   C. Some gymnast names are spelled differently in the two data frames.

   D. `left_join()` only works on data frames read in from a CSV file, not on `tribble()`s.

**1b.**
Fix the code: join `paris_scores` and `noc_codes`, in that order, with `left_join()`, telling R which columns to match on.
Save the result as `scores_codes`.

::: callout-tip
When the key columns have different names in the two data frames, specify them with `join_by()`, for example `join_by(id == identifier)`, where the first name is the column in the left data frame and the second name is the column in the right data frame.
:::

### Question 2

Join the two data frames with an `inner_join()`, in both orders:

**2a.**
`paris_scores` first, then `noc_codes`.
Save the result as `inner_scores_first`.

**2b.**
`noc_codes` first, then `paris_scores`.
Save the result as `inner_codes_first`.

Look carefully at both results.
How do they compare in number of rows, number of columns, and column order?

::: {.callout-tip title="Render, commit, and push"}
Render, commit, and push your changes to GitHub with the commit message "Added answers for Questions 1 and 2".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 3

Join the two data frames with an `anti_join()`, in both orders:

**3a.**
`paris_scores` first, then `noc_codes`.
Save the result as `anti_scores_first`.

**3b.**
`noc_codes` first, then `paris_scores`.
Save the result as `anti_codes_first`.

**3c.**
Your ultimate goal is to construct a dataset where you have all gymnasts who are in the `paris_scores` data frame along with their event scores and the country they competed for.
In order to accomplish this, you will need to look up some gymnasts' NOC codes in an alternate resource, since they're not all in the `noc_codes` data frame.
Record in `q3_which_join` the join function you would use to find out *whose* NOC codes you need to look up:

   A. `left_join()`

   B. `inner_join()`

   C. `anti_join()`

   D. `full_join()`

### Question 4

The `paris_scores` data frame is in a *wide* format: each apparatus has its own column.
Reshape it into a *long* format with one row per gymnast-apparatus combination, using `pivot_longer()`.
The result should have exactly three columns, `gymnast`, `event`, and `score`, in that order, where `event` contains the former column names (`balance_beam` or `floor_exercise`) and `score` contains the scores.
Save the result as `scores_long`.

::: {.callout-tip title="Render, commit, and push"}
Render, commit, and push your changes to GitHub with the commit message "Added answers for Questions 3 and 4".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

## Part 2

**Paris 2024 routines**

For Part 2 of this lab, you'll work with a new dataset, `gymnastics`, which contains information about the routines performed by gymnasts in the women's artistic gymnastics competition at the 2024 Paris Olympics.
**Each row represents a single routine**, so a gymnast who competed on all four apparatus in qualification appears in at least four rows.

The template reads the data into R with:

```{r}
#| eval: false
gymnastics <- read_csv("data/gymnastics.csv")
```

The variables in this dataset are:

| Variable       | Description                                                           |
|----------------|-----------------------------------------------------------------------|
| `name`         | Gymnast's name, in "Given Surname" order                              |
| `country`      | Country name                                                          |
| `country_code` | 3-letter National Olympic Committee (NOC) code                        |
| `event`        | Apparatus: Vault, Uneven Bars, Balance Beam, Floor Exercise           |
| `competition`  | Which round: Qualification, Team Final, All-Around Final, Event Final |
| `d_score`      | Difficulty score (a number greater than zero, in increments of 0.1)   |
| `e_score`      | Execution score (a number between 0 and 10)                           |
| `penalty`      | Neutral deduction                                                     |
| `score`        | Routine total = `d_score` + `e_score` - `penalty`                     |
| `rank`         | Placement for that row's competition                                  |

### Question 5

Get to know the data.
`glimpse()` at the `gymnastics` data frame, then record:

-   `n_routines`: the number of routines (rows) in the data frame.
-   `n_gymnasts`: the number of distinct gymnasts in the data frame.

Write code that computes each of these numbers rather than typing them in.

::: callout-tip
The `distinct()` function keeps one row for each distinct value of the variable(s) you give it, dropping the duplicates.
So the number of distinct gymnasts is the number of rows in the result of `distinct(name)`:

```{r}
#| eval: false
gymnastics |>
  distinct(name) |>
  nrow()
```
:::

### Question 6

Olympic order is the specific order in which gymnasts compete the apparatus.
Read about Olympic order in [this NBC Olympics explainer](https://www.nbcolympics.com/news/gymnastics-101-olympic-competition-format).

**6a.**
Based on the article, record the Olympic order of the four events in women's artistic gymnastics in `olympic_order`, as a character vector with the event names spelled exactly as they appear in the data, e.g., `c("Event 1", "Event 2", "Event 3", "Event 4")`.

**6b.**
Run the following code in your Console and analyze the result.

```{r}
#| eval: false
gymnastics |>
  pull(event) |>
  as.factor() |>
  levels()
```

Record in `q6_why_alpha` the option that best explains why the events are listed in this order in the output:

   A. When a character variable is converted to a factor, R sorts the levels alphabetically by default.

   B. The levels are listed in the order in which each event first appears in the data.

   C. The levels are listed in Olympic order.

   D. The levels are listed in order of the average score on each event.

**6c.**
Write code to change the `event` variable so that its levels reflect Olympic order and save the new variable, with the same name, in the `gymnastics` data frame, i.e., overwrite the `gymnastics` data frame with the new version of the `event` variable.
Then, display the new ordering of the levels with `levels(gymnastics$event)`.

::: callout-tip
`fct_relevel()` takes a factor (or a character variable) and a vector of level names, and returns a factor with the levels in that order.
You already have that vector from part (a).
Use it inside `mutate()`.
:::

::: {.callout-tip title="Render, commit, and push"}
Render, commit, and push your changes to GitHub with the commit message "Added answers for Questions 5 and 6".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 7

In the All-Around Final, each gymnast competes on all four apparatus and her four scores are added up.
The `gymnastics` data frame has one row per routine, so each All-Around finalist appears in four rows.
Reshape the data so that each finalist appears in **one** row, with one column per apparatus.

**7a.**
In a single pipeline, keep only the All-Around Final routines, keep only the `name`, `country`, `event`, and `score` columns, and then use `pivot_wider()` to spread the events into columns.
The result should have one row per gymnast and the columns `name`, `country`, `Vault`, `Uneven Bars`, `Balance Beam`, and `Floor Exercise`.
Save the result as `aa_wide`.

::: callout-tip
Dropping the other columns *before* pivoting matters: `pivot_wider()` uses every column you do not pivot to identify a row, so if columns like `rank` or `d_score` are still present, each routine remains its own row.
:::

**7b.**
Starting from `aa_wide`, add a column called `total` containing the sum of the four apparatus scores, and arrange the rows in descending order of `total`.
Save the result as `aa_totals`.

::: callout-tip
Column names that contain spaces are not "proper" names in R, so to refer to them you need to surround them with backticks, e.g. `` `Uneven Bars` ``.
:::

**7c.**
Record the winner's total score in `q7_winner_total`, as a number.
If everything worked, the order of your rows should match the official All-Around Final results.

### Question 8

**8a.**
Create a vector called `gymnasts_of_interest` containing the names of two to five All-Around finalists whose scores you want to visualize.
For example:

```{r}
#| eval: false
gymnasts_of_interest <- c("Gymnast One", "Gymnast Two")
```

Make sure that the spelling of the names matches how they appear in the data.

**8b.**
In a single pipeline, filter `gymnastics` to include only the All-Around Final routines of your `gymnasts_of_interest`, and save the result as `aa_selected`.
Each gymnast should appear in exactly four rows.

**8c.**
Using `aa_selected`, create a plot of score (y-axis) vs. event (x-axis) for these gymnasts, saved as `aa_plot`.

Requirements for the plot:

-   Data should be represented with points as well as lines connecting each gymnast's four scores.
-   Each gymnast should be represented by a different color line and different color and shape points.
-   Axes and legend should be properly labeled.
-   The plot should have an appropriate title (and optionally a subtitle).
-   The plot should be customized in at least one way -- you could use a different than default theme, a different than default color scale, or some other customization.

Because you reordered the levels of `event` in Question 6, the apparatus will appear on the x-axis in Olympic order.

::: callout-tip
When the x-axis variable is categorical, `geom_line()` needs to be told which points belong together on the same line.
Add `group = name` to the aesthetics of the line layer: `geom_line(aes(group = name))`.
:::

::: {.callout-tip title="Render, commit, and push"}
Render, commit, and push your changes to GitHub with the commit message "Added answers for Questions 7 and 8".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

## Part 3

**Continents**

To qualify for major competitions such as the Olympics, gymnasts must compete in their Continental Championship: the Pan American Games, European Championships, Asian Championships, African Championships, or Oceania Championships.
One question of interest, then, is whether gymnasts from different continents score differently.

The `continents` dataset contains countries and the continent each one is in.
The template reads it in with `continents <- read_csv("data/continents.csv")`.
Its variables are:

| Variable    | Description                                         |
|-------------|-----------------------------------------------------|
| `entity`    | Country name                                        |
| `code`      | Country code (3-letter ISO code for most countries) |
| `year`      | Year the continent assignment was recorded          |
| `continent` | Continent the country is in                         |

::: callout-note
The Pan American Games feature athletes from both North and South American countries.
For this analysis, you don't have to combine the two continents.
:::

### Question 9

**9a.**
Join the two data frames based on the `country` variable in `gymnastics` and the `entity` variable in `continents`, and save the result as `gym_continents`.
Make sure that your joined data frame contains **all** rows of the `gymnastics` data frame, whether or not the country was found in `continents`.

**9b.**
In a single pipeline starting from `gym_continents`, produce a summary of the average vault score by continent.
The result should have two columns, `continent` and `mean_vault`, and one row per continent.
Save it as `vault_by_continent`.

**9c.**
Record the continent with the highest average vault score in `q9_highest` and the continent with the lowest in `q9_lowest`, as strings spelled exactly as in the data.
Ignore the row where `continent` is `NA`.

**9d.**
Your summary has a row where `continent` is `NA`: some countries in `gymnastics` did not find a match in `continents`.
Using `anti_join()` followed by `distinct(country)`, write code to identify which countries in `gymnastics` have no match in `continents`.
Save the result as `unmatched`.

**9e.**
Look up each unmatched country in both data frames (the data viewer, or `filter()`, will help).
Two of them fail to match because the two data frames use different names for the same country.
Record in `q9_reason` the one that fails for a **different** reason:

   A. Great Britain

   B. Chinese Taipei

   C. AUT

::: {.callout-tip title="Render, commit, and push"}
Render, commit, and push your final changes to GitHub with a meaningful commit message.

Make sure to commit and push all changed files so that your Git pane is empty afterward.
We keep repeating this because it's important and because we see students forget to do this.
:::

# Wrap-up

## Submission

::: callout-warning
Before you wrap up the assignment, make sure all of your files are updated in your GitHub repo.
We will be checking these to make sure you have been committing and pushing changes (with descriptive commit messages).

You must turn in **both** the `.qmd` file and the PDF to Gradescope by the submission deadline to be considered "on time".
:::

Download your completed `lab-3.qmd` and `lab-3.pdf` files.
To download a file from Posit Cloud, go to the **Files** pane, select the file by marking its **checkbox**, click **More**, then **Export**.

Submit **BOTH** `lab-3.qmd` and `lab-3.pdf` in Gradescope (accessed via Canvas), due by the deadline. (You will NOT need to select pages for the PDF.)

::: callout-important
## Checklist

Make sure you have:

-   answered all questions, using the exact object names given in each question
-   run the check chunk for each question and reviewed the results
-   rendered your Quarto document
-   committed and pushed everything to your GitHub repository so that the Git pane in RStudio is empty
-   downloaded your QMD and PDF files from RStudio onto your computer
-   uploaded your `lab-3.qmd` **and** `lab-3.pdf` files to the Lab 3 Gradescope assignment and confirmed the autograder ran successfully
:::
